---
title: "Lab 03 - Nobel laureates"
author: "Haley Lam"
date: "Jan 30, 2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

```{r}
nrow(nobel)
ncol(nobel)
head(nobel)
```

The nobel dataset contains `r nrow(nobel)` observations and `r ncol(nobel)` variables.
Each row represents one nobel prize winner.

### Exercise 2

```{r}
nobel_living <- nobel %>%
  filter((is.na(died_date)), 
         (!is.na(country)),
         (gender != "org")
  )
```

There are now `r nrow(nobel_living)` observations of living nobel laureates.

### Exercise 3

```{r}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )
```

```{r}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

```{r}
ggplot(data = nobel_living_science, aes(x = country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip()
```


Plot for category of a prize and whether they were in US.

Based on this, the Buzzfeed article is right, that over half were based in the US when they won the nobel prize.


### Exercise 4
```{r}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = ifelse(born_country_original == "USA", "USA", "Other")
  )
```

```{r}
summary(nobel_living_science$born_country_us == "USA")
```

105 were born in the USA.

### Exercise 5
```{r}
ggplot(data = nobel_living_science, aes(x = country_us, fill = born_country_us)) +
  geom_bar() +
  facet_wrap(~ category) +
  coord_flip()
```

This shows that a significant amount of people who won nobel prizes in the US were born outside of the US. This lines up with the visualizations in the BuzzFeed article.

### Exercise 6
```{r}
nobel_living_science %>%
  filter(country_us == "USA", born_country_us == "Other") %>%
  count(born_country, sort = TRUE)

```

Germany and UK are the most common. It's different from the Buzzfeed article presumably because this is more updated?